---
title: "Analysis"
author: "Sam Milnes"
date: "3/15/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Unemployment Rate = x(independent), market = y(dependent)

```{r}
#attach(S_PDataFINAL)
#Data Every month from 3/1/2001 to 5/1/2003

Mode = function(x, na.rm = FALSE){
  if(na.rm){
    x = x[!is.na(x)]
  }
  ux = unique(x)
  return(ux[which.max(tabulate(match(x, ux)))])
}
f = table(S_PDataFINAL$OpenPrice)
which(f==max(f))

```

```{r}

SP_mean = mean(S_PDataFINAL$OpenPrice)
SP_median = median(S_PDataFINAL$OpenPrice)
SP_mode = Mode(S_PDataFINAL$OpenPrice)
SP_standardDeviation = sd(S_PDataFINAL$OpenPrice)
SP_variance = var(S_PDataFINAL$OpenPrice)

hist(S_PDataFINAL$OpenPrice,
     main = "Histogram for S&P 500 Open Price",
     xlab = "S&P 500 Price",
     freq = F,
     col = 'green4', border = 'black',
     breaks = 10,
     bg = "grey80")
       
lines(density(S_PDataFINAL$OpenPrice))



boxplot(S_PDataFINAL$OpenPrice,
        names = c(" S&P 500 Open Price"),
        show.names = TRUE, main = "Boxplot for S&P 500 Open Price", 
        col = 'green', border = 'blue',
        ylab = 'Price  ($)')

plot(x = S_PDataFINAL$OpenPrice, ylab = "S&P 500 Price($)",
     main = "Line Graph - S&P 500 Price ($)",
     type = 'l', col = 'green')
#The x variable in this line graph is just the order of the data. 
plot(x = S_PDataFINAL$UnemploymentRate, ylab = "Unemployment Rate (%)",
     main = "Line Graph - US Unemployment",
     type = 'l', col = 'green')
#The x variable in this line graph is just the order of the data. 

```

```{r}
plot(S_PDataFINAL$UnemploymentRate,S_PDataFINAL$OpenPrice,
     main = "S&P 500 Price vs US Unemployment Rate Scatterplot",
     xlab = "US Unemployment Rate (%)",
     ylab = "S&P 500 Price ($)",
     pch = 20,
     col = 'green4',
     bg = "grey80")sur
abline(lm(S_PDataFINAL$OpenPrice ~ S_PDataFINAL$UnemploymentRate), 
       col = "black",)
legend("topright", legend = c("y = -103.58x + 2318.49"), bty = 'n')

LinearModelUnemploymentPrice = lm(S_PDataFINAL$OpenPrice ~ S_PDataFINAL$UnemploymentRate)
summary(LinearModelUnemploymentPrice) #gives values for liner equation
#y = (slope)* x + y-intercept
#y = -103.58x + 2318.49 <- estimated regression line equation

```


Approximate the dependent variable (y) value by choosing a value for the independent variable (x)
x = 2
```{r}
approximateY = -103.58*2 + 2318.49
```
When the unemployment rate is 2%, we can get an approximate value of 2111.33 for the S&P500 price.


y = 1000

y = -103.58x + 2318.49 -> 1000 = -103.58x + 2318.49 -> -1318.49 = -103.58x -> x = -1318.49 / -103.58
```{r}
approximateX = -1318.49 / -103.58
```
When the S&P price is 1000, we can get an approximate value of 12.7% for the unemployment rate.

```{r}
r_value = cor(S_PDataFINAL$UnemploymentRate,S_PDataFINAL$OpenPrice)
```

We get a negative correlation coefficient, which for our data means that our two variables,
x and y, are moving in opposite directions of each other.
More or less it means that the variables are an inverse correlation, they do not move in the same direction.
Suggests that there is a weak negative correlation between S&P500 Price (Y) and Unemployment Rate (X)

```{r}
rSquared_value = (r_value * r_value)
```
From our coefficient of determination, we can see that almost 8% of our dependent
variable is caused or directly related to our independent variable.
Accounts for the variability in our y-variable that is attributed to the relationship determined by our
linear model between our x-variable and y-variable.

8% of the variability in the s&P 500 can be attributed to the linear model relationship between
the unemployment rate and the S&P 500 price.


```{r}
qqnorm(S_PDataFINAL$OpenPrice, pch =20, frame = FALSE, col = "green4", main = "Normal QQ-Plot for S&P 500 Price")

#Adding a reference line
qqline(S_PDataFINAL$OpenPrice, col = "black", lwd = 1)
```
From the normal QQ plot, we can see that the dependent variable is overall,
normal distributed with some outlying points.
skews to the left first(pulls below) and then skews to the right(pulls above)



#95% Confidence level - Create Margin of Error for our dependent variable (Y)
```{r}
price_sample_sizeN = nrow(S_PDataFINAL) #Finds population size
price_population_meanX = mean(S_PDataFINAL$OpenPrice) #Finds population mean

price_sample_var = var(S_PDataFINAL$OpenPrice) #Finds the sample variance
price_sample_sd = sd(S_PDataFINAL$OpenPrice) #Finds the sample standard deviation

#For Z, we need POPULATION var/sd, so we will switch using formulas
price_population_var = (((price_sample_sizeN - 1) / price_sample_sizeN) * price_sample_var)
price_population_sd = sqrt(price_population_var)

#Find the Z-Critial Values
price_alpha_95 = 0.05 #95% = (1-ALPHA) * 100%
price_z_CV_95 = qnorm(1-price_alpha_95/2)

#Creating Margin of Error
price_z_ME_95 = price_z_CV_95 * (price_population_sd / sqrt(price_sample_sizeN))

#Finds 95% CI
price_lower_bound_z_95 = price_population_meanX - price_z_ME_95 
price_upper_bound_z_95 = price_population_meanX + price_z_ME_95
cat("The 95% confidence interval for the true population mean price of S&P500 is: (",
    price_lower_bound_z_95, "," , price_upper_bound_z_95,").")
```
When we have a 95% confidence level, the length of the interval is around 181.2.
We are 95% confident that the population S&P500 Price is between the interval of 1596.3 and 1777.5.


```{r}
#95% Confidence level - Create Margin of Error for our dependent variable (Y)
unemployment_sample_sizeN = nrow(S_PDataFINAL) #Finds population size
unemployment_population_meanX = mean(S_PDataFINAL$UnemploymentRate) #Finds population mean

unemployment_sample_var = var(S_PDataFINAL$UnemploymentRate) #Finds the sample variance
unemployment_sample_sd = sd(S_PDataFINAL$UnemploymentRate) #Finds the sample standard deviation

#For Z, we need POPULATION var/sd, so we will switch using formulas
unemployment_population_var = (((unemployment_sample_sizeN - 1) / unemployment_sample_sizeN) * unemployment_sample_var)
unemployment_population_sd = sqrt(unemployment_population_var)

#Find the Z-Critial Values
unemployment_alpha_95 = 0.05 #95% = (1-ALPHA) * 100%
unemployment_z_CV_95 = qnorm(1-unemployment_alpha_95/2)

#Creating Margin of Error
unemployment_z_ME_95 = unemployment_z_CV_95 * (unemployment_population_sd / sqrt(unemployment_sample_sizeN))

#Finds 95% CI
unemployment_lower_bound_z_95 = unemployment_population_meanX - unemployment_z_ME_95 
unemployment_upper_bound_z_95 = unemployment_population_meanX + unemployment_z_ME_95
cat("The 95% confidence interval for the true population mean unemployment rate is: (",
    unemployment_lower_bound_z_95, "," , unemployment_upper_bound_z_95,").")
```
When we have a 95% confidence level, the length of the interval is around .5
We are 95% confident that the population unemployment rate is between the interval of 5.8 and 6.3.





